---
title       : Commuting Methods by State, Income
subtitle    : Creating an interactive Shiny app using 2014 ACS data
author      : Matthew Sedlar
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [bootstrap, mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## What is the ACS?

ACS stands for the American Community Survey.

From the dataset abstract:

> An annual nationwide survey that collects information such as age, race, income, commute time to work, home value, veteran status, and other data.

The data is open to the public and available through [Data.Gov](http://www.data.gov/). 

This particular application uses the [Means of Transportation to Work by Selected Characteristics for Workplace Geography](http://factfinder.census.gov/bkmk/table/1.0/en/ACS/14_1YR/S0804/0100000US.04000) 2014 American Community Survey 1-Year Estimates from the U.S. Census Bureau.

---

## Why Commuting Methods by Income Groups?

* The dataset includes statistics on everything from commuting method by age, sex, race to even citizenship status. 
* Since owning a vehicle is often considered a luxury item, I was curious about whether workers 16 years and over in lower income groups relied more on public transportation and carpooling while workers in higher income groups drove alone to work. But obviously it varies state by state, because some states have better public transportation options than others.
* Why not build an app that helps conduct an exploratory analysis of the data?

---

## What the Data Looks Like

```{r echo=FALSE, message=FALSE, cache=FALSE}

library(dplyr)

# unzipping data
unzip("data/ACS_14_1YR_S0804.zip",exdir="data/")

data <- read.csv("data/ACS_14_1YR_S0804.csv", skip = 1, nrows=1)

tidydata <- data

# selecting only columns necessary for calculations
tidydata <- tidydata %>% select(Geography, #1
                                Car..truck..or.van....drove.alone..Estimate..EARNINGS.IN.THE.PAST.12.MONTHS..IN.2014.INFLATION.ADJUSTED.DOLLARS..FOR.WORKERS...Workers.16.years.and.over.with.earnings, #2
                                Car..truck..or.van....drove.alone..Estimate..EARNINGS.IN.THE.PAST.12.MONTHS..IN.2014.INFLATION.ADJUSTED.DOLLARS..FOR.WORKERS...Workers.16.years.and.over.with.earnings....1.to..9.999.or.loss, #3
                                                                Car..truck..or.van....drove.alone..Estimate..EARNINGS.IN.THE.PAST.12.MONTHS..IN.2014.INFLATION.ADJUSTED.DOLLARS..FOR.WORKERS...Workers.16.years.and.over.with.earnings....10.000.to..14.999, #4
                                                                Car..truck..or.van....drove.alone..Estimate..EARNINGS.IN.THE.PAST.12.MONTHS..IN.2014.INFLATION.ADJUSTED.DOLLARS..FOR.WORKERS...Workers.16.years.and.over.with.earnings....15.000.to..24.999, #5
                                                                Car..truck..or.van....drove.alone..Estimate..EARNINGS.IN.THE.PAST.12.MONTHS..IN.2014.INFLATION.ADJUSTED.DOLLARS..FOR.WORKERS...Workers.16.years.and.over.with.earnings....25.000.to..34.999, #6
                                Car..truck..or.van....drove.alone..Estimate..EARNINGS.IN.THE.PAST.12.MONTHS..IN.2014.INFLATION.ADJUSTED.DOLLARS..FOR.WORKERS...Workers.16.years.and.over.with.earnings....35.000.to..49.999, #7
                                                                Car..truck..or.van....drove.alone..Estimate..EARNINGS.IN.THE.PAST.12.MONTHS..IN.2014.INFLATION.ADJUSTED.DOLLARS..FOR.WORKERS...Workers.16.years.and.over.with.earnings....50.000.to..64.999, #8
                                                                Car..truck..or.van....drove.alone..Estimate..EARNINGS.IN.THE.PAST.12.MONTHS..IN.2014.INFLATION.ADJUSTED.DOLLARS..FOR.WORKERS...Workers.16.years.and.over.with.earnings....65.000.to..74.999, #9
                                                                Car..truck..or.van....drove.alone..Estimate..EARNINGS.IN.THE.PAST.12.MONTHS..IN.2014.INFLATION.ADJUSTED.DOLLARS..FOR.WORKERS...Workers.16.years.and.over.with.earnings....75.000.or.more) #10
                                
# convert percentages to numeric, percentage by decimal
tidydata[3:10] <- sapply(tidydata[3:10],function(x) as.numeric(x)/100)

colnames(tidydata) <- c("State",
                         "Total Drove Alone",
                         "1 to 9,999 or loss",
                         "10,000 to.14.999",
                         "15,000 to 24,999",
                         "25,000 to 34,999",
                         "35,000 to 49,999",
                         "50,000 to 64,999",
                         "65,000 to 74,999",
                         "75,000 or more")

tidydata <- t(tidydata)
colnames(tidydata) <- "Observation"

```

Below is a data frame (transposed with margins of error excluded to fit the slide) of the state of Alabama and the estimated percentage of people who drove alone by income levels.

```{r echo=FALSE, cache=FALSE}
library(knitr)

kable(tidydata)
```

---

## How the App Works

The user inputs state and income group and the app calculates the following:

* T = Estimated total of workers 16 years and older for commuting method and income group.
* I = Percentage (in decimals) of workers in selected income group
* N = Number of estimated workers for commuting method.
* d = drive alone, c = carpooled, p = public transportation.
* S = Overall sample of workers

Example of workers who drove alone:
$T_d=N_d(I_d)$

Upper and Lower Margins: $N_d(I_d \pm Margin)$

$S=T_d+T_c+T_p$

Each bar in the chart represents: $\frac{T_d}{S},\frac{T_c}{S},\frac{T_p}{S}$

---

## App Output

```{r}

```

---
